---
title: "Movie Data Analysis and Movie Rating Prediction"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


This is a Data analysis document on Movies Data using statistical methods and Predictive algorithms

```{r cars, message=FALSE, warning=FALSE}
library(caret)
library(e1071)
library(ggplot2)
library(dplyr)
```

#Reading the input file
```{r cars2, message=FALSE, warning=FALSE}
movies <- read.csv("C:/Users/dell/Downloads/movies.csv")
str(movies)
```

# Removing rows with blank values

```{r pressure, echo=F}
print(paste('Total number of rows ',nrow(movies),sep = ':'))
```
```{r pressure2, echo=T}
gotna <- which(!complete.cases(movies))
movies <- movies[-gotna,]
```
```{r pressure3, echo=F}
print(paste('Total number of rows ',nrow(movies),sep = ':'))
```

#Histogram of Number of reviews
```{r pressure4, echo=T}
ggplot(aes(x = critics_score), data = movies) + geom_histogram(bins = 20, color = 'white') + ggtitle('Histogram of Number of reviews')
summary(movies$critics_score)
```


#Histogram of Movie Ratings
```{r pressure5, echo=T}
ggplot(aes(x = imdb_rating), data = movies) + geom_histogram(bins = 20, color = 'white') + ggtitle('Histogram of Scores')
summary(movies$imdb_rating)
```


#Histogram of Movie Release Year
```{r pressure6, echo=T}
ggplot(aes(x = thtr_rel_year), data = movies) + geom_histogram(color='white') +
  ggtitle('Histogram of Theater Release Year')
```


#Boxplot of Movie Ratings and Release Year
```{r pressure7, echo=T}
boxplot(imdb_rating ~ thtr_rel_year, data=movies, col='indianred')
title("IMDB rating vs Release year")
```

#Which genre have the highest scores?
```{r pressure8, echo=T}
genre_group <- group_by(movies, genre)
movie_by_genre <- summarise(genre_group,
                              mean_score = mean(imdb_rating),
                              n = n()) 

ggplot(aes(x = genre, y = n, fill = genre), data = movie_by_genre) + geom_bar(stat = 'identity') + theme(legend.position = "none", axis.text=element_text(size=6)) +
  coord_flip() + ggtitle('Genre vs Number of Movies')
```


#Which Directors movie has the highest scores?
```{r pressure9, echo=T}
director_group <- group_by(movies, director, genre)
movie_by_director <- summarise(director_group,
                               mean_score = mean(imdb_rating))

movie_by_director <- movie_by_director
movie_by_director <- movie_by_director[with(movie_by_director, order(-mean_score)), ]
movie_by_director <- head(movie_by_director, 20)
ggplot(aes(x = mean_score, y = director), data = movie_by_director) +
  geom_point(aes(color = genre), size = 2) + xlab("Mean Score") + 
  ylab("Director Name")+ theme_minimal() + ggtitle('Director, Genres & Scores')
```

#Spliting the data into train and test for a predictive model
```{r pressure10, echo=T}
set.seed(2017)
train_size <- 0.8 
train_index <- sample.int(length(movies$imdb_rating), length(movies$imdb_rating) * train_size)
train_sample <- movies[train_index,]
test_sample <- movies[-train_index,]
```

#Building a Muliple Linear Regression model using selected Features
```{r pressure11, echo=T}
fit <- lm(imdb_rating ~ critics_score + runtime  + thtr_rel_month + thtr_rel_day + thtr_rel_year + imdb_num_votes + audience_score, data=train_sample)
summary(fit)
```

#Checking the Accuracy of a model on train and test data
```{r pressure12, echo=T}
train_sample$pred_score <- predict(fit, train_sample)
train_corr <- (cor(train_sample$pred_score, train_sample$imdb_rating))
train_mae <- (mean(abs(train_sample$pred_score - train_sample$imdb_rating)))
c(train_corr^2, train_mae)

test_sample$pred_score <- predict(fit, test_sample)
test_corr <- (cor(test_sample$pred_score, test_sample$imdb_rating))
test_mae <- (mean(abs(test_sample$pred_score - test_sample$imdb_rating)))
c(test_corr^2, test_mae)
```

#Conclusion
The most important factor that affect movie rating is the runtime and the critic score.
